### 2008-03-21 trunk/Papers/dl-gre-08/gre.tex
minor edits

In the examples above, it is because $\mathsf{flower} \sqcap \exists
\mathsf{in}. \mathsf{hat}$ denotes exactly $\{f_2\}$ that we can say
``the flower in the hat'' to refer to $f_2$.  This perspective
provides a general framework into which many existing GRE approaches
fit: Traditional attribute selection \cite{Dale1995} corresponds to
building purely DL formulas that are conjunctions of atoms; relational
REs as in \newcite{dale91:_gener_refer_expres_invol_relat} are
formulas of \el; and so on.  We will pursue the idea of organizing GRE
approaches with respect to the variant of DL they use further in
Section~\ref{sec:related}.

For the rest of this paper, we assume that we are generating a
singular RE, i.e.\ the target referent $A$ will generally be a
singleton set.  In this case, we will only be able to generate a
formula that denotes exactly $A = \{a\}$ (i.e., a RE that uniquely
refers to $a$) if there is no other individual $b$ to which $a$ is
similar; otherwise, any formula that is satisfied by $a$ is also
satisfied by $b$.  Conversely, if we know that $a$ is not similar to
any other individual, then there is a formula that is satisfied by $a$
and not by anything else; this formula can serve as a unique singular
RE.  In other words, we can reduce the $\gL$-GRE problem for a given
model to the problem of computing the $\gL$-similarity sets of this
model.

@@

The \alc\ algorithm is capable of computing the \alc-similarity sets
of the model in time $O(nm)$, where $n$ is the number of individuals
in the domain, and $m$ is the number of pairs of individuals that are
related by some relation.  However, it will freely introduce negations
in the case distinctions, which can make the resulting formula hard to
realize (see also Section~\ref{sec:discussion-realization}).  This is
why we also present an algorithm for the \el-similarity sets; \el\
corresponds to positive relational REs, which are generally much
easier to realize.

@@

The \el\ algorithm has a harder time with the example in
Fig.~\ref{fig:dale-haddock}b \cite{Stone1998a}.  While it will
correctly identify $r_1$ as ``the rabbit in the hat'' and $f_2$ as
``the flower in the bathtub'', it will not be able to compute a RE for
$f_1$ because $f_2$ is \el-similar to $f_1$.  Indeed, the algorithm
terminates with $\RE$ containing both $\mathsf{flower}$ and
$\mathsf{flower} \sqcap \exists \mathsf{in}.\mathsf{hat}$.  This is a
typical pattern for asymmetrical cases of similarity in \el: If there
are two formulas $\varphi_1$ and $\varphi_2$ in the output set with
$\interp{\varphi_1} \subseteq \interp{\varphi_2}$, then there is
generally some individual $b \in \interp{\varphi_2} -
\interp{\varphi_1}$ such that all individuals in $\interp{\varphi_1}$
are similar to $b$, but not vice versa.  By contrast, the \alc\
algorithm can exploit the greater expressivity of \alc\ to split
$\mathsf{flower}$ into the two new formulas $\mathsf{flower} \sqcap
\exists \mathsf{in}.\mathsf{hat}$ and $\mathsf{flower} \sqcap \neg
\exists \mathsf{in}.\mathsf{hat}$, generating a unique RE for $f_1$ as
well.

>>>>

In the examples above, it is because $\mathsf{flower} \sqcap \exists
\mathsf{in}. \mathsf{hat}$ denotes exactly $\{f_2\}$ that we can say
``the flower in the hat'' to refer to $f_2$.  This perspective
provides a general framework into which many existing GRE approaches
fit: Traditional attribute selection \cite{Dale1995} corresponds to
building DL formulas that are conjunctions of atoms; relational REs as
in \newcite{dale91:_gener_refer_expres_invol_relat} are formulas of
\el; and so on.  We will pursue the idea of organizing GRE approaches
with respect to the variant of DL they use further in
Section~\ref{sec:related}.

For the rest of this paper, we assume that we are generating a
singular RE, i.e.\ the target referent $A$ will be a singleton set.
In this case, we will only be able to generate a formula that denotes
exactly $A = \{a\}$ (i.e., a RE that uniquely refers to $a$) if there
is no other individual $b$ to which $a$ is similar; otherwise, any
formula that is satisfied by $a$ is also satisfied by $b$.
Conversely, if we know that $a$ is not similar to any other
individual, then there is a formula that is satisfied by $a$ and not
by anything else; this formula can serve as a unique singular RE.  In
other words, we can reduce the $\gL$-GRE problem for a given model to
the problem of computing the $\gL$-similarity sets of this model.

@@

The \alc\ algorithm computes the \alc-similarity sets of the model in
time $O(nm)$, where $n$ is the number of individuals in the domain,
and $m$ is the number of pairs of individuals that are related by some
relation.  However, it will freely introduce negations in the case
distinctions, which can make the resulting formula hard to realize
(see also Section~\ref{sec:discussion-realization}).  This is why we
also present an algorithm for the \el-similarity sets; \el\
corresponds to positive relational REs, which are generally much
easier to realize.

@@

The \el\ algorithm has a harder time with the example in
Fig.~\ref{fig:dale-haddock}b \cite{Stone1998a}.  While it will
correctly identify $r_1$ as ``the rabbit in the hat'' and $f_2$ as
``the flower in the bathtub'', it will not be able to compute a RE for
$f_1$ because $f_1$ is \el-similar to $f_2$.  Indeed, the algorithm
terminates with $\RE$ containing both $\mathsf{flower}$ and
$\mathsf{flower} \sqcap \exists \mathsf{in}.\mathsf{hat}$.  This is a
typical pattern for asymmetrical cases of similarity in \el: If there
are two formulas $\varphi_1$ and $\varphi_2$ in the output set with
$\interp{\varphi_1} \subseteq \interp{\varphi_2}$, then there is
generally some individual $b \in \interp{\varphi_2} -
\interp{\varphi_1}$ such that all individuals in $\interp{\varphi_1}$
are similar to $b$, but not vice versa.  By contrast, the \alc\
algorithm can exploit the greater expressivity of \alc\ to split
$\mathsf{flower}$ into the two new formulas $\mathsf{flower} \sqcap
\exists \mathsf{in}.\mathsf{hat}$ and $\mathsf{flower} \sqcap \neg
\exists \mathsf{in}.\mathsf{hat}$, generating a unique RE for $f_1$ as
well.
